Choosing the Choice: Reflections on Modelling Decisions and Behaviour in Demographic Agent-Based Models

### 2017-04-28 article.tex
TRACE comment addressed

This iterative approach places considerable demands on the modeller in terms of developing the model, but also in exercising their judgement as to what should be included, and identifying where a component should be iterated out. This is however an expensive exercise, both because augmenting a model is time consuming, and because evaluating complex models poses considerable challenges. The challenge of evaluation can be somewhat ameliorated by the use of techniques like sensitivity analysis (see \citet{Thiele2014} for a review of several techniques, and \citet{Oakley2004} for an alternative approach), which can support the modeller in identifying how the individual parameters of a model contribute to overall outputs. In this context, sensitivity analysis refers to a broad range of techniques that enable understanding, how the various model inputs affect the output (result) variables, and how important the model parameters are relative to one another \citep{Oakley2004, Thiele2014}. One important benefit to sensitivity analysis is that parameters which do not significantly impact the results of a model can be removed. This too has limitations, because as we discuss later in relation to heterogeneity in agents (section \ref{heterogeneity-in-decision-strategies}), context can take the form of both process, and parameters, and sensitivity analysis is directly informative only about the latter.As a result, the onus remains on the modeller to use their best judgement. The application of such judgement can be documented, however: \cite{Schmolke2010} suggest that developmental cycles are recorded in a standardised format, the TRACE framework, so the process of rejection of alternatives through which the final model came about can be reconstructed.

>>>>

This iterative approach places considerable demands on the modeller in terms of developing the model, but also in exercising their judgement as to what should be included, and identifying where a component should be iterated out. This is however an expensive exercise, both because augmenting a model is time consuming, and because evaluating complex models poses considerable challenges. The challenge of evaluation can be somewhat ameliorated by the use of techniques like sensitivity analysis (see \citet{Thiele2014} for a review of several techniques, and \citet{Oakley2004} for an alternative approach), which can support the modeller in identifying how the individual parameters of a model contribute to overall outputs. In this context, sensitivity analysis refers to a broad range of techniques that enable understanding, how the various model inputs affect the output (result) variables, and how important the model parameters are relative to one another \citep{Oakley2004, Thiele2014}. One important benefit to sensitivity analysis is that parameters which do not significantly impact the results of a model can be removed. This too has limitations, because as we discuss later in relation to heterogeneity in agents (section \ref{heterogeneity-in-decision-strategies}), context can take the form of both process, and parameters, and sensitivity analysis is directly informative only about the latter. As a result, the onus remains on the modeller to use their best judgement.

The application of judgement in the process of model development should be documented, however: \cite{Schmolke2010} suggest that developmental cycles are recorded in a standardised format, the Transparent and Comprehensive Framework for Ecological Modelling (TRACE) framework. This framework sets down a structure in which the model development, testing and application phases of a project can be described to the reader, so that is possible for the reader to reconstruct the process through which the final model was chosen. The authors suggest that TRACE documentation should be a living document resembling a lab notebook, which also facilitates interaction with decision-makers. Such documentation of the model development process provides transparency about the decisions made and the justifications for doing so.
